Propositions as Sessions, Mechanically

### 2015-04-10 paper/background.tex
Commuting conversion example (Wadler's style). Fixed small issues.

\citeauthor{Wadler:2014}'s recent work~\cite{Wadler:2014} brings together a
process calculus, CP, based on the earlier work by
\citeauthor{Caires:2010:STI}, and a functional language, GV, based on LAST by
\citeauthor{Gay:2010:LAST}. Figure~\ref{fig:gvtnts} presents the types and
terms GV with types ranged over by $T$, session types ranged over by $S$ and
terms ranged over by $L$, $M$ and $N$. GV session types are a modified version
of the definition presented in \S~\ref{sec:pre} where \lstinline{end} has been
replaced with $\outend$ and $\inend$. The intention is that $\outend$
($\inend$) terminates a session type whose last operation is an output
(input), and they correspond to the units of the respective linear logical
connectives \todo{Do I need to elaborate on the translation? Yet another
  figure?}. CP provides linear logical connectives with session type
interpretations inspired by
\citeauthor{Caires:2010:STI}~\cite{Caires:2010:STI} shown in
Figure~\ref{fig:cptnts} along with the terms of CP. The process calculus CP is
slightly different to $\pi$DILL by \citeauthor{Caires:2010:STI}, apart from
being classical in nature, the constructs for server replication are modified:
CP provides separate weakening and contraction rules whereas these are encoded
in one rule for process reduction in $\pi$DILL. Additionally, rather than a
two-sided sequent calculus, \citeauthor{Wadler:2014} uses one-sided sequents
for defining CP, which provides a more intuitive presentation of session type
duality and a clearer correspondence to a $\lambda$-calculi surface syntax
like GV. \citeauthor{Wadler:2014} describes a continuation-passing style (CPS)
translation from GV types and terms to CP classical linear logic propositions
and processes, respectively. As presented, GV does not support all features of
CP and thus the translation is one way (GV to
CP). \citeauthor{Lindley:2014:SAP}~\cite{Lindley:2014:SAP} extend GV to
provide support for polymorphism and replication following a similar CPS
translation scheme. Lastly,
\citeauthor{Lindley:2014:SPS}~\cite{Lindley:2014:SPS} describe operational
semantics for GV and their system has greater similarities with LAST than
prior GV works, for example, by treating send and receive as linear functions
rather than language primitives.

>>>>

\citeauthor{Wadler:2014}'s recent work~\cite{Wadler:2014} brings together a
process calculus, CP, based on the earlier work by
\citeauthor{Caires:2010:STI}, and a functional language, GV, based on LAST by
\citeauthor{Gay:2010:LAST}. Figure~\ref{fig:gvtnts} presents the types and
terms GV with types ranged over by $T$, session types ranged over by $S$ and
terms ranged over by $L$, $M$ and $N$. GV session types are a modified version
of the definition presented in \S~\ref{sec:pre} where \lstinline{end} has been
replaced with $\outend$ and $\inend$. The intention is that $\outend$
($\inend$) terminates a session type whose last operation is an output
(input), and they correspond to the units of the respective linear logical
connectives (see \S~\ref{sec:trans}). CP provides linear logical connectives
with session type interpretations shown in Figure~\ref{fig:cptnts} along with
the terms of CP.

The process calculus CP is slightly different to $\pi$DILL by
\citeauthor{Caires:2010:STI}, besides from being classical (rather than
intuitionistic), the constructs for server replication are modified: CP
provides separate weakening and contraction rules whereas these are encoded in
one rule for process reduction in $\pi$DILL. Additionally, rather than a
two-sided sequent calculus, \citeauthor{Wadler:2014} uses one-sided sequents
for defining CP, which provides a more intuitive presentation of session type
duality and a clearer correspondence to a $\lambda$-calculi surface syntax
like GV.

\input{ccoutput}

Subject reduction in CP proceeds by cut reductions of the logical operators
against their duals and \textit{commuting conversions}. Commuting conversions
push a cut inside a logical operator and the intuition is that they permit
further communication to occur which was guarded by a cut. An example of a
commuting conversion on output is shown in Figure~\ref{fig:ccoutput}. On the
left-hand side $R$ can interact with its environment, but commuting the output
and the cut prevents $R$ from doing this. As explained by
\citeauthor{Wadler:2014}, the semantics of the rule are either $x$ is bound by
an outer cut and will be eliminated eventually by cut reduction, or the
process is blocked waiting for external communication on channel $x$. This
differs from the lambda calculus where the evaluation strategy is
deterministic (e.g. left-to-right call-by-value semantics).

\citeauthor{Wadler:2014} describes a continuation-passing style (CPS)
translation from GV types and terms to CP classical linear logic propositions
and processes, respectively. Freedom from deadlock within GV is assured
through the translation of CP and the top-level cut elimination property of CP
(see \S~\ref{sec:cutelim}).

\citeauthor{Wadler:2014}'s GV does not support all features of CP and thus the
translation is one way (GV to
CP). \citeauthor{Lindley:2014:SAP}~\cite{Lindley:2014:SAP} extend GV to
provide support for polymorphism and replication following a similar CPS
translation scheme. Lastly,
\citeauthor{Lindley:2014:SPS}~\cite{Lindley:2014:SPS} describe operational
semantics for GV and their system has greater similarities with LAST than
prior GV works, for example, by treating send and receive as linear functions
rather than language primitives.

### 2015-04-10 paper/formalisation.tex
Commuting conversion example (Wadler's style). Fixed small issues.

While this diverges with the linear logic rule for top, I believe it to be
sound within the CP/GV eco-system such that this variant of CP still leads to
a valid translation to GV. In a private communication, Philip Wadler echoed
these remarks but pointed out that this is likely more to do with GV
\todo{what's another way to say ``too small'' vs. ``too large''}.

@@

\subsubsection{Top-level Cut Elimination}

@@

\subsection{CPS translation from GV to CP}

>>>>

While this diverges with the linear logic rule for top, I believe this variant
of CP still admits a valid translation from GV. In a private communication,
Philip Wadler echoed these remarks but pointed out that this is likely because
the GV-to-CP translation does not require the full expressiveness of CP.

@@

\subsubsection{Top-level Cut Elimination}\label{sec:cutelim}

@@

\subsection{CPS translation from GV to CP}\label{sec:trans}
